Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-09-22 Paper/app.tex
..

The active subspace was computed using the iterative procedure outlined in Algorithm~\ref{alg:grad}.
The convergence of the eigenvectors
was examined by tracking $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$, plotted in 
Figure~\ref{fig:conv_app} (right). In Figure~\ref{fig:conv_app} (left), individual
components of the converged eigenvector are illustrated. The convergence was established using
a $\tau$ value of 0.02. Finite difference was used to compute model gradients at 40 samples in
the input domain. Hence, a total of 1360 model evaluations were required for this purpose.  
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.8\textwidth]{./Figures/eigv10}
\caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector,
obtained using the perturbation-based strategy. Right: The quantity,  
$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
is plotted for successive iterations to illustrate the convergence behavior.}
\label{fig:conv_app}
\end{center}
\end{figure}
%

In Figure~\ref{fig:hd} (left), we illustrate the resulting eigenvalue spectrum. The second eigenvalue is
found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace
is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right), generated
using both, perturbation-based and regression-based strategies. 
%
\begin{figure}[htbp]
 \begin{center}
   \includegraphics[width=0.45\textwidth]{./Figures/eig_33D}
   \includegraphics[width=0.45\textwidth]{./Figures/ssp_33D}
\caption{Left: Eigenvalue spectrum of $\hat{\mat{C}}$. Right: SSP of the computed active subspace, and the
corresponding 1-D linear fit.} 
\label{fig:hd}
\end{center}
\end{figure}
%
Specifically
in the case of perturbation-based strategy, the SSP is observed to exhibit a linear variation that
is reasonably
captured by the corresponding straight-line fit, $\tilde{G}$, constructed using the sequence of steps
outlined in Section~\ref{sub:ac}.
On the other hand, the SSP constructed using the regression-based strategy
is relatively more scattered. Consequently, the two surrogates (straight-line fits) are observed to exhibit a
discrepancy unlike our earlier observation for the 19-dimensional problem in Figure~\ref{fig:comp_ssp}.

\subsection{Surrogate Verification}
\label{sub:verify}

The 1-dimensional surrogate ($\tilde{G}$) shown in Figure~\ref{fig:hd}~(right) is investigated
for its accuracy as well as the ability to capture the uncertainty in the
model output in two ways. Firstly, we estimate the relative L-2 norm of the discrepancy~($\varepsilon_d$)
between estimates of ignition delay in the case of H$_2$/O$_2$ reaction, obtained using the 
model output and the surrogate in the following equation:
%
\be
\varepsilon_d = \frac{\|G(\bm{\xi}) - \tilde{G}(\bm{\xi})\|_2}{\|G(\bm{\xi})\|_2}
\ee
%
Model evaluations ($G$) at 10$^{4}$ samples in the input domain were used to evaluate $\varepsilon_d$. 
It was estimated to be 1.35$\times10^{-2}$ and 1.02$\times10^{-1}$
using $\tilde{G}$ based on the perturbation-based and the regression-based strategies respectively. Hence,
$\varepsilon_d$ was found to increase by an order of magnitude when using the regression-based strategy.
Therefore, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained
using the perturbation-based approach is remarkably accurate with a relative error of less than 1.5$\%$, whereas
the regression-based approach resulted in a relative error of about 10$\%$.

@@

The normalized activity scores ($\tilde{\nu}_{i,r}$) based on the dominant eigenspace, obtained
using the two strategies (perturbation-based and regression-based) are compared with the 
total Sobol' indices
in Figure~\ref{fig:as_33D}. Note that the Sobol' indices were computed using the verified
1-dimensional surrogate ($\tilde{G}$) in the active subspace, computed using the 
perturbation-based strategy. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new}
\caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
\label{fig:as_33D}
\end{center}
\end{figure}
%
Several useful inferences can be drawn from the above plot. Firstly, the total Sobol indices are found to be
consistent
with the normalized activity scores evaluated using the perturbation-based strategy. The regression-based
strategy yields consistent results for the $A_i$'s, but it does not capture the sensitivity associated with
the most important parameter, $E_{a,15}$. This is because numerical errors are incurred when approximating
the model gradients using a linear regression fit in this case. 
Secondly, our findings indicate that   
the ignition delay is predominantly sensitive towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
moderately sensitive towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
found to be low or negligible. 
%Note that this observation can also be exploited for reducing the 
%dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.
%The sensitivity-driven dimension reduction for uncertainty quantification has been
%discussed in our earlier effort~\cite{Vohra:2018}. In this work, however, we focus on dimension reduction using the active 
%subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. Furthermore, we have
%shown that the dominant eigenspace can be used to approximate global sensitivity measures with no additional effort.

>>>>

The active subspace was computed using the iterative procedure outlined in Algorithm~\ref{alg:grad}.
The convergence of the eigenvectors
was examined by tracking $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$, plotted in 
Figure~\ref{fig:conv_app} (right). In Figure~\ref{fig:conv_app} (left), individual
components of the converged eigenvector that constitutes the active subspace,
are illustrated. The convergence was established using
a $\tau$ value of 0.02. Finite difference was used to compute model gradients at 40 samples in
the input domain. Hence, a total of 1360 model evaluations were required for this purpose.  
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.8\textwidth]{./Figures/eigv10}
\caption{Left: Individual components of the dominant eigenvector in the
active subspace,
obtained using the perturbation-based strategy. Right: The quantity,  
$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
is plotted for successive iterations to illustrate the convergence behavior.}
\label{fig:conv_app}
\end{center}
\end{figure}
%

In Figure~\ref{fig:hd} (left), we illustrate converged estimates of the eigenvalues.
 The second eigenvalue is
found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace
is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right), generated
using both, perturbation-based and regression-based strategies. 
%
\begin{figure}[htbp]
 \begin{center}
   \includegraphics[width=0.45\textwidth]{./Figures/eig_33D}
   \includegraphics[width=0.45\textwidth]{./Figures/ssp_33D}
\caption{Left: Eigenvalue spectrum of $\hat{\mat{C}}$. Right: SSP of the computed active subspace, and the
corresponding 1-D linear fit.} 
\label{fig:hd}
\end{center}
\end{figure}
%
Specifically
in the case of perturbation-based strategy, the SSP is observed to exhibit a linear variation that
is reasonably
captured by the corresponding straight-line fit, $\tilde{G}$, constructed using the sequence of steps
outlined in Section~\ref{sub:ac}.
On the other hand, the SSP constructed using the regression-based strategy
is relatively more scattered. Consequently, the two surrogates (straight-line fits) are observed to exhibit a
discrepancy unlike our earlier observation for the 19-dimensional problem in Figure~\ref{fig:comp_ssp}.

\subsection{Surrogate Assessment}
\label{sub:verify}

The 1-dimensional surrogate ($\tilde{G}$) shown in Figure~\ref{fig:hd}~(right) is investigated
for its accuracy as well as the ability to capture the uncertainty in the
model output in two ways. Firstly, we estimate the relative L$_2$ norm of the discrepancy~($\varepsilon_d$)
between estimates of ignition delay in the case of H$_2$/O$_2$ reaction, obtained using the 
model output and the surrogate in the following equation:
%
\be
\varepsilon_d = \frac{\|G(\bm{\xi}) - \tilde{G}(\bm{\xi})\|_2}{\|G(\bm{\xi})\|_2}
\ee
%
Model evaluations ($G$) at 10$^{4}$ samples in the input domain were used to evaluate $\varepsilon_d$. 
It was estimated to be 1.35$\times10^{-2}$ and 1.02$\times10^{-1}$
using $\tilde{G}$ based on the perturbation-based and the regression-based strategies respectively. Hence,
$\varepsilon_d$ was found to increase by an order of magnitude when using the regression-based strategy.
Therefore, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained
using the perturbation-based approach is remarkably accurate with a relative error of less than 1.5$\%$, whereas
the regression-based approach resulted in a relative error of about 10$\%$.

@@

The normalized activity scores ($\tilde{\nu}_{i,r}$) based on the dominant eigenspace, obtained
using the two strategies (perturbation-based and regression-based) are compared with the 
total Sobol' indices
in Figure~\ref{fig:as_33D}. Note that the Sobol' indices were computed using the verified
1-dimensional surrogate ($\tilde{G}$) in the active subspace, computed using the 
perturbation-based strategy. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new}
\caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
\label{fig:as_33D}
\end{center}
\end{figure}
%
Several useful inferences can be drawn. Firstly, the total Sobol indices are found to be
consistent
with the normalized activity scores evaluated using the perturbation-based strategy. The regression-based
strategy yields consistent results for the $A_i$'s, but it does not capture the sensitivity associated with one of the most important parameters, $E_{a,15}$. This is because numerical errors are incurred when approximating
the model gradients using a linear regression fit in this case. 
Secondly, our findings indicate that   
the ignition delay is predominantly sensitive towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
moderately sensitive towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
found to be low or negligible. 
%Note that this observation can also be exploited for reducing the 
%dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.
%The sensitivity-driven dimension reduction for uncertainty quantification has been
%discussed in our earlier effort~\cite{Vohra:2018}. In this work, however, we focus on dimension reduction using the active 
%subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. Furthermore, we have
%shown that the dominant eigenspace can be used to approximate global sensitivity measures with no additional effort.

### 2018-09-22 Paper/conc.tex
..

Additionally, the surrogates were assessed by comparing probability distributions (based on kernel density estimation)
for the ignition delay, constructed using surrogate predictions and model evaluations at the same set of 10$^4$
random samples. The PDFs based on model evaluations and the 1-dimensional surrogate from the perturbation-based
strategy were found to be almost identical. However, the PDF based on 1-dimensional surrogate from the
regression-based strategy although captured the modal estimate with reasonable accuracy, was observed to
under-estimate the uncertainty in the ignition delay. Mean estimates of the ignition delay using model evaluations
and predictions using the two surrogates at 10$^4$ samples were found to be in agreement. However, as expected, the
standard deviation, computed using predictions from the surrogate based on the regression-based strategy was
found to be accurate only upto the first significant digit. Moreover, it was observed that the activity scores,
computed using the regression-based strategy failed to capture the sensitivity towards the 
most important parameter, $E_{a,15}$. 
We attribute this discrepancy to two potential sources of numerical errors: (1) Regression-based approximation of the
gradient of the model output, and (2) Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
error incurred by the linear regression fit in this case. Based on our findings in this work, it can be said that
the regression-based strategy would be useful in situations involving complex mechanisms with a large number of
uncertain inputs thereby rendering the UQ analysis intensive, and in applications where the focus is on a
reasonable approximation of the statistics of the QoI.

 The authors would like to highlight that the computational framework presented in this work is agnostic to the choice
 of the chemical system and could thus be easily adapted for other systems. However, it is important to note that
 its applicability depends upon the nature of the relationship between the model output and the inputs. Specifically,
 the model output is required to be differentiable throughout the considered input domain to be able to evaluate
 the active subspace with reasonable accuracy. Hence, it is a realistic possibility that an active subspace might not
 exist for a given input-output relationship. While the active subspace could be exploited for a forward propagation of
 the uncertainty from inputs to the output, the low-dimensional surrogate could further be used to accelerate 
 calibration of the rate-controlling parameters in a Bayesian setting. Moreover, the activity scores could help
 identify important stages in a complex reaction mechanism and thereby guide development of a simplified
 mechanism. Furthermore, dimension reduction using active subspaces could be useful in studies aimed at
 investigating the inadequacy or model error associated with the reaction rate laws for such systems.

>>>>

Additionally, the surrogates were assessed by comparing probability distributions (based on kernel density estimation)
for the ignition delay, constructed using surrogate predictions and model evaluations at the same set of 10$^4$
random samples. The PDFs based on model evaluations and the 1-dimensional surrogate from the perturbation-based
strategy were found to be almost identical. However, the PDF based on 1-dimensional surrogate from the
regression-based strategy although captured the modal estimate with reasonable accuracy, was observed to
under-estimate the uncertainty in the ignition delay. Mean estimates of the ignition delay using model evaluations
and predictions using the two surrogates at 10$^4$ samples were found to be in agreement. However, as expected, the
standard deviation, computed using predictions from the surrogate based on the regression-based strategy was
found to be accurate only upto the first significant digit. Moreover, it was observed that the activity scores,
computed using the regression-based strategy failed to capture the sensitivity towards the 
most important parameter, $E_{a,15}$. 
We attribute this discrepancy to two potential sources of numerical errors: (1) Regression-based approximation of the
gradient of the model output, and (2) Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
error incurred by the linear regression fit in this case. Hence, the regression-based approximation of the gradient 
of the model output although reduces computational effort, the present investigations clearly reveal a potential pitfall associated
with this approach. Therefore, this approach should only be preferred in situations involving intensive simulations, and 
where the focus is essentially on estimating the statistics of the QoI as opposed to a more detailed analysis such as
parametric sensitivity analysis.

 We would like to highlight that the computational framework presented in this work is agnostic to the choice
 of the chemical system and could thus be easily adapted for other systems as long as the quantity of interest
is continuously differentiable in the considered domain of the inputs.
We have demonstrated that the active subspace could be exploited for a forward propagation of
 the uncertainty from inputs to the output. However, the activity scores and the low-dimensional surrogate could further guide optimal allocation of computational resources for  
 calibration of the important rate-controlling parameters in a Bayesian setting. 
Additionally, dimension reduction using active subspaces could assist in developing robust formulations
for predicting discrepancy between simulations and measurements due to epistemic uncertainty in the model inputs.

### 2018-09-22 Paper/cover.tex
..
minor ...

\begin{center}
\textsc{Efficient UQ for chemical systems using active subspaces}

@@

\begin{tabbing}
Corresponding Author: \hspace{5mm} \= Sankaran Mahadevan\\
       \>  Department of Civil and Environmental Engineering\\
       \>  Vanderbilt University\\
       \>  272 Jacobs Hall, VU Mailbox: PMB 351831 \\
       \>  Nashville, TN 37235 \\
       \> \\
Phone: \> (615) 322-3040 \\
Fax:   \> (615) 343-3773 \\
Email: \>  sankaran.mahadevan@vanderbilt.edu   \\
\\
Submitted to: \> \textit{The Journal of Chemical Physics} \\
\>  August 2018\\

>>>>

\begin{center}
%\textsc{Efficient UQ for chemical kinetics using active subspaces}
\textsc{Subspace-based dimension reduction for chemical kinetics applications with epistemic uncertainty}

@@

\begin{tabbing}
Corresponding Author: \hspace{5mm} \= Manav Vohra\\
       \>  Department of Civil and Environmental Engineering\\
       \>  Vanderbilt University\\
       \>  279 Jacobs Hall, VU Mailbox: PMB 351831 \\
       \>  Nashville, TN 37235 \\
       \> \\
Phone: \> (615) 322-3040 \\
Fax:   \> (615) 343-3773 \\
Email: \>  manav.vohra@vanderbilt.edu   \\
\\
Submitted to: \> \textit{The Journal of Chemical Physics} \\
\>  September 2018\\

### 2018-09-22 Paper/method.tex
..

As discussed earlier, gradient estimation using finite differences in the 
perturbation-based strategy requires additional model evaluations at the
neighboring points in the input domain. 
Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
model evaluations are needed. On the other hand, the regression-based
strategy uses the available set of $N$ model evaluations to approximate
the gradient. The computational effort is therefore 
reduced by a factor $(d+1)$ in this case. The specific sequence of steps
for computing the active subspace is discussed as follows.

@@

	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_k$ 
		\State Compute $\bm{g}^i = \nabla_{\bm{\xi}_i}f(\bm{\xi}_i)$, 
             	$i=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
		\State Compute $\hat{\mat{C}}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$
		\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)}
		 \hat{\mat{W}}^{(k)\top}$
		\State Partition the eigenspace of $\hat{\mat{C}}$ as shown in Step 7
		\State Compute $\delta \hat{\mat{W}}_{1,j}^{(k)}$ = 
                       \scalebox{1.25}{$\frac{\|(\hat{\mat{W}}_{1,j}^{k})^2 - 
                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|}$}, 
                       $j = 1,\ldots,r$.
		\If {$\max\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \Lambda_j w_{i,j}^2$,
	$i=1,\ldots,N_p$.
	\State Normalize $\nu_{i,r}(f)$ as $\tilde{\nu}_{i,r}(f)$ = \scalebox{1.25}{$\frac{\nu_{i,r}(f)}{\sum_i\nu_{i,r}(f)}$}.

@@

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem with uncertain $A_i$'s. For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum ($\lambda_1, \ldots, 
\lambda_{19}$) with the same using a much 
smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
$N$=1000 and $n$ = $\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Since 
$\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we expect 
a 1-dimensional active subspace. Hence, the first eigenvector 
sufficiently captures the uncertainty in
the  model output. To further confirm this, we evaluate a relative L-2 norm of the difference
 ($\varepsilon_\text{L-2}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000
and $n$~=~$\{20,40,80,120\}$.  The quantity, $\varepsilon_\text{L-2}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ for $n$ = 20, 40, and 80; and $\mathcal{O}(10^{-3})$ for $n$ = 120.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational advantage.

@@

As mentioned earlier, the model output $f(\bm{\xi})$ predominantly varies in the active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the two strategies is 
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the 
perturbation-based and the regression-based strategies for computing the active subspace.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the response in ignition
delay based on the considered prior marginals for $A_i$'s is observed to be approximately linear in the
active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.

>>>>

As discussed earlier, gradient estimation using finite differences in the 
perturbation-based strategy requires additional model evaluations at the
neighboring points in the input domain. 
Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
model evaluations are needed. On the other hand, gradient estimation
using the regression-based approach involves a series of linear regression
fits to subsets of available evaluations as discussed 
in~\cite{Constantine:2015}. Hence, the computational effort is  
reduced by a factor $(d+1)$ when using the regression-based approach.
The specific sequence of steps
for computing the active subspace is discussed as follows.

@@

	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_k$ 
		\State Compute $\bm{g}^i = \nabla_{\bm{\xi}_i}f(\bm{\xi}_i)$, 
             	$i=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
		\State Compute $\hat{\mat{C}}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$
		\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)}
		 \hat{\mat{W}}^{(k)\top}$
		\State Partition the eigenspace of $\hat{\mat{C}}$ as shown in Step 7
		\State Compute $\delta \hat{\mat{W}}_{1,j}^{(k)}$ = 
                       \scalebox{1.25}{$\frac{\|(\hat{\mat{W}}_{1,j}^{k})^2 - 
                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|}$}, 
                       $j = 1,\ldots,r$.
		\If {$\max\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \lambda_j w_{i,j}^2$,
	$i=1,\ldots,N_p$.
	\State Normalize $\nu_{i,r}(f)$ as $\tilde{\nu}_{i,r}(f)$ = \scalebox{1.25}{$\frac{\nu_{i,r}(f)}{\sum_i\nu_{i,r}(f)}$}.

@@

To assess its feasibility and suitability, we implement
Algorithm~\ref{alg:grad} to compute the active subspace for 
the 19-dimensional H$_2$/O$_2$ reaction kinetics
problem with uncertain $A_i$'s. For the purpose of verification,
$\hat{\mat{C}}$ was initially constructed using a large set of samples
($N$~=~1000) in the input domain. The gradient was estimated using
finite difference, and hence, a total of 20,000 model runs were performed. 
In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
the resulting normalized eigenvalue spectrum ($\lambda_1, \ldots, 
\lambda_{19}$) with the same using a much 
smaller set of samples, $n$~=~$\{20,40,80,120\}$.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
\caption{A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
obtained using a much larger sample size, $N$~=~1000. 
%Right: Relative $L_2$ 
%norm of the difference between
%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
$N$=1000 and $n$ = $\{20,40,80,120\}$ samples.
} 
\label{fig:eig_comp}
\end{center}
\end{figure}
%
We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
are approximated 
reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
to improve with the sample size. Since 
$\lambda_1$ is roughly three orders of magnitude larger than $\lambda_2$, we expect 
a 1-dimensional active subspace. Hence, the first eigenvector 
sufficiently captures the uncertainty in
the  model output. To further confirm this, we evaluate a relative L$^2$ norm of the difference
 ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
%
\be
\varepsilon_{\text{L}^{2}}^{N-n} = 
\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
\ee
%
The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
$\mathcal{O}(10^{-2})$ for $n$ = 20, 40, and 80; and $\mathcal{O}(10^{-3})$ for $n$ = 120.
Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
reasonable accuracy in this case. 
The iterative strategy therefore offers a significant potential for computational gains.

@@

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the igntion delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in the active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the two strategies is 
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the 
perturbation-based and the regression-based strategies for computing the active subspace.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $A_i$'s is observed to be approximately linear in the
active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.
